---
title: "proposal"
author: "ShufanXia"
date: "10/16/2021"
output: html_document
---

```{r}
#load library
library(tseries)
library(forecast)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(GGally)
#detach("package:Metrics", unload=TRUE)
```

```{r}


csv2020<- read.csv("Data/1090006_36.01_-78.94_tmy-2020.csv",stringsAsFactors = F)#,header= FALSE)
basicinfo <- csv2020[1,]
data<-csv2020[c(-1,-2),seq(1,14,1)]
data<-mutate_all(data, function(x) as.numeric(as.character(x)))
colnames(data)<- csv2020[2,seq(1,14,1)]

data$date<-paste(paste("2020",data$Month,data$Day,sep="-"), paste(data$Hour,data$Minute,sep=":"))
data$time <-strptime(data$date,format="%Y-%m-%d  %H:%M")
data[,seq(1,14,1)]<- sapply(data[,seq(1,14,1)],as.numeric)
tsGHI<-ts(data$GHI)#,frequency=24*365)
tsTemp<-ts(data$Temperature,frequency=24*365)

tsdata <- lapply(data[,seq(6,14,1)],function(t) ts(t,frequency=24*365))

GHIhead <- window(tsGHI,start=c(1,1),end=c(1,8040))
GHItail <- window(tsGHI,start=c(1,8040),end=c(1,8760))
```

stationary?:
the exact time does not matter, the relative distribution same whenever you look at the dist
This is not stationary

Trend: parabolic, peak during summer
Seasonal: yes: both annual seasonal and 24hour seasonal
fixed frequency of up and down: 24 hours, dirunal cycle 
(is this seasonal or cyclic, i think there is no cyclic behavior)

frequency=8766 or 1???

```{r}
GHIdf <- data.frame(GHI = tsGHI, hour = seq(1,length(tsGHI),1))
ggplot(data=GHIdf[seq(12,8760,24),],aes(x=hour,y=GHI))+geom_point()+labs(x="GHI at noon over a year")

ts.plot(tsGHI,col="red3",gpars=list(xlab="Hour since Jan.1st 00:30", ylab="Hourly GHI", lty=c(1:3)))
ts.plot(tsGHI[1:2000],col="red3",gpars=list(xlab="hour", ylab="GHI", lty=c(1:3)))
ts.plot(tsGHI[1:120],col="red3",gpars=list(xlab="hour", ylab="GHI", lty=c(1:3)))
ts.plot(tsGHI[1200:1300],col="red3",gpars=list(xlab="Hour since Jan.1st 00:30", ylab="Hourly GHI", lty=c(1:3)))
```
```{r}
#just to have seasonal plot
seasonaldata <- aggregate(data[,c("DHI","DNI","GHI","Dew Point","Surface Albedo","Wind Direction","Wind Speed","Temperature" ,"Pressure")],list(data$Month,data$Hour),mean)

seasonMeanGHIts <- ts(seasonaldata$GHI,frequency = 24)
ggseasonplot(seasonMeanGHIts,polar=TRUE)+ylab("Mean Hourly GHI")+labs(title="Hourly GHI averaged by month")+scale_color_discrete(name= "Month")
```

#lag plot
An elliptical plot suggests that the data comes from a single-cycle sinusoidal model.
Strongest correlation at lag=1 and lag =2

```{r}
gglagplot(tsGHI)

```



```{r}
p1<-autoplot(tsGHI)+labs(y="GHI")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-1.5,0) , units = "lines" ) )
p11 <-autoplot(tsdata[['DNI']])+labs(y="GNI")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-1.5,0) , units = "lines" ) )
p12 <-autoplot(tsdata[['DHI']])+labs(y="GHI")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-1.5,0) , units = "lines" ) )

grid.arrange(p1, p11,p12,ncol = 1)

p1<-autoplot(tsGHI)+labs(y="GHI")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-0.5,0) , units = "lines" ) )
p2<-autoplot(tsTemp)+labs(y="Temp")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-0.5,0) , units = "lines" ) )
p3 <-autoplot(tsdata[['Surface Albedo']])+labs(y="Surface Albedo")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-0.6,0) , units = "lines" ) )
p4 <-autoplot(tsdata[['Dew Point']])+labs(y="Dew Point")+theme(plot.margin = unit( c(0,0,-0.5,0) , units = "lines" ) )
grid.arrange(p1, p2,p3,p4,ncol = 1)

p5 <-autoplot(tsdata[['Wind Direction']])+labs(y="Wind Direction")+theme(axis.title.x = element_blank(),plot.margin = unit( c(0,0,-0.5,0) , units = "lines" ) )
p6 <-autoplot(tsdata[['Wind Speed']])+labs(y="Wind Speed")+theme(plot.margin = unit( c(0,0,-0.5,0) , units = "lines" ) )



grid.arrange(p5,p6,ncol = 1)
```

```{r}
qplot(data$DHI,data$GHI)+ylab("GHI (W/m^2)")+xlab("DHI")
qplot(data$DNI,data$GHI)+ylab("GHI (W/m^2)")+xlab("DNI")
qplot(data$Temperature,data$GHI)+ylab("GHI (W/m^2)")+xlab("Temperature (Celsius)")
qplot(data$`Dew Point`,data$GHI)+ylab("GHI (W/m^2)")+xlab("Dew Point")
qplot(data$`Wind Direction`,data$GHI)+ylab("GHI (W/m^2)")+xlab("Wind Direction")
qplot(data$`Wind Speed`,data$GHI)+ylab("GHI (W/m^2)")+xlab("Wind Speed")
qplot(data$Pressure,data$GHI)+ylab("GHI (W/m^2)")+xlab("Pressure")
qplot(data$`Surface Albedo`, data$GHI) +ylab("GHI (W/m^2)")+xlab("Surface Albedo")#+ geom_point(alpha = .5,colour="blue4") #+  geom_smooth(method = (y~x))


```

#let's make surface albdeo categorical??
#how about Pressure?
```{r}
hist(data$`Surface Albedo`)
hist(data$`Surface Albedo`[data$`Surface Albedo`<0.8])
length(data$`Surface Albedo`[data$`Surface Albedo`>0.8])
```
# predictor slection here

```{r}
data$SurfaceAlbedoLevel[data$`Surface Albedo`==0.11]<-"0.11"
data$SurfaceAlbedoLevel[data$`Surface Albedo`==0.12]<-"0.12"
data$SurfaceAlbedoLevel[data$`Surface Albedo`==0.13]<-"0.13"
data$SurfaceAlbedoLevel[data$`Surface Albedo`==0.14]<-"0.14"
data$SurfaceAlbedoLevel[data$`Surface Albedo`==0.15]<-"0.15"
data$SurfaceAlbedoLevel[data$`Surface Albedo`==0.87]<-"0.87"
data$SurfaceAlbedoLevel <- as.factor(data$SurfaceAlbedoLevel)
data$hour_since <- seq(1,8760,1)

linearmodel1 <- lm(GHI~Temperature+SurfaceAlbedoLevel+Pressure+`Dew Point`+`Wind Speed`+`Wind Direction`,data=data)
summary(linearmodel1)
plot(linearmodel1)
ggplot(data, aes(x=as.numeric(hour_since), y=linearmodel1$residual)) +
  geom_point(alpha = .5,colour="blue4") +
  geom_hline(yintercept=0,col="red3") + labs(title="Residuals vs Time")

linearmodel2 <- lm(GHI~Temperature+SurfaceAlbedoLevel+Pressure+`Dew Point`+`Wind Speed`+`Wind Direction`+Hour+Month,data=data)
summary(linearmodel2)
plot(linearmodel2)

ggplot(data, aes(x=as.numeric(time), y=linearmodel2$residual)) +
  geom_point(alpha = .5,colour="blue4") +
  geom_hline(yintercept=0,col="red3") + labs(title="Residuals vs Time")

# trend
```
# predictor selection
```{r}
```

Stationary test and autocorrelation test?
Augmented Dickey-Fuller Test reject null hypothesis: so this is stationary
KPSS Test for Level Stationarity: the null hypothesis is actually "stationary", so reject the null hypothesis

Two tests say different things,
(Hyndman, R.J) This is not staionary becasue correlatoin decreases only gradually
Paritial correlation after 1 lag is still significant, linear regression model is not sufficient.

The decreasing in the ACF as the lags increase is due to the trend, while the sinusoidal shape is due the durinal cycle

Partial autocorrelation plot suggests p=4 become small enough: ar p=4
Don't see how MA would help


```{r}

tsGHI %>% ggtsdisplay(lag.max=80,main="Hourly GHI")
adf_test <- adf.test(tsGHI,alternative = 'stationary')
print(adf_test)

kpss_test <- kpss.test(tsGHI)
print(kpss_test)


acf(tsGHI[1:8760],lag.max=80)
pacf(tsGHI[1:8760],lag.max=80)
```

First order differenceing makes it stationary
from autocorrelation: dies off at lag = 4 --> ma= 4??
from partial autocorrelation: dies off at 2, but comes back again
```{r}
tsDiffOne <- diff(tsGHI)

autoplot(tsDiffOne)
tsDiffOne%>%ggtsdisplay()

adf_test <- adf.test(tsDiffOne,alternative = 'stationary')
print(adf_test)

kpss_test <- kpss.test(tsDiffOne)
print(kpss_test)


acf(tsDiffOne,lag.max=200)
pacf(tsDiffOne,lag.max=100)
# if only ar, lag=29 or more
# if I add auto
```
```{r}
diff(tsGHI,24)%>%diff() ->   tsDiff24

tsDiff24%>%ggtsdisplay()

adf_test <- adf.test(tsDiff24,alternative = 'stationary')
print(adf_test)

kpss_test <- kpss.test(tsDiff24)
print(kpss_test)


acf(tsDiff24,lag.max=80,main="Seasonal differenceing with lag=24")
pacf(tsDiff24,lag.max=80,main="Seasonal differenceing with lag=24")
```
```{r}
ggpairs(data[,c("GHI","Dew Point","Surface Albedo","Wind Direction","Wind Speed","Temperature","Pressure")])
```

# let's fit an AR model
```{r}
# autoARmodel <- auto.arima(tsGHI,d=1,max.q=0)
# autoARmodel
armodel <- arima(tsGHI,order=c(26,1,0))#method"CSS"cross validation
armodel
```
# check resiudal (model validation)
```{r}
checkresiduals(armodel)
acf(armodel$residuals,lag.max=80)
pacf(armodel$residuals,lag.max=80)
```
# compute accuracy
MASE = 1.50
```{r}
# refitARmodel <- arima(GHIhead,c(26,1,0))
# fcastAR <- forecast(refitARmodel,h=720)
# autoplot(GHIhead) +
#   autolayer(fcastAR, series="forecast", PI=FALSE)
# fcastAR %>% autoplot()
# accuracy(fcastAR,GHItail)
```


# let's use arima model 
select based on AIC
R estimates the ARIMA model, it uses maximum likelihood estimation (MLE)

d = 1: confirm that 1 differencing makes the series stationary
p = 4, q=2

They indicate the possible ARIMA (p, 1, q) model, where, usually, p = 1–5, q = 1–3, and d is usually 1 [42]
# I could try forcing q=0, and it will  an AR model
# i need to add MA part
```{r}
autoarimaModel1<-auto.arima(GHIhead)
autoarimaModel1
# fit <- Arima(tsGHI,order=c(4,1,2))
# fit
```

# aesseing the model
shows that all autocorrelations are not within the threshold limits, indicating that the residuals are not behaving like white noise. 
A portmanteau test returns a small p-value, also suggesting that the residuals are not white noise.

```{r}
checkresiduals(autoarimaModel1)
acf(residuals(autoarimaModel1))
pacf(residuals(autoarimaModel1))
```


#accuracy
use (p,d,q)=(4,1,2) to refit arima model and forecast on the last month
MASE = 2.18
# let's make test dta for last month
```{r}
arimamodel_refit <- arima(GHIhead,order=c(4,1,2))
arimamodel_refit
arimafcast <- forecast(arimamodel_refit,h=2190)
autoplot(GHIhead) +
  autolayer(arimafcast, series="forecast", PI=FALSE)

arimafcast %>% autoplot()
accuracy(arimafcast,GHItail)

```


#This is bad, I think clearly it says seasonal. I think it calls for a seasonal model
#auto arima (d=0) model on the first-order differencing series, giving the same result: confirmed
```{r}
arma1<-auto.arima(tsDiffOne,d=0)#,order=c(4,2))
arma1
autoarimaModel1
checkresiduals(arma1)
```

```{r}
arima27 <- arima(tsGHI,order =c(26,1,2))
```
```{r}
checkresiduals(arima27 )
acf(arima27 $residuals,lag.max=80)
pacf(arima27 $residuals,lag.max=80)
```
```{r}
arima27_refit <- arima(GHIhead,order=c(26,1,2))
arima27_refit
```
```{r}
checkresiduals(arima27_refit)
acf(arima27_refit $residuals,lag.max=80)
pacf(arima27_refit  $residuals,lag.max=80)
```

#ar_arimafcast <- forecast(ar_arima_refit,h=2190)
```{r}
autoplot(GHIhead) +
  autolayer(ar_arimafcast, series="forecast", PI=FALSE)
ar_arimafcast %>% autoplot()
accuracy(ar_arimafcast,GHItail)

```



# let's throw in some predictor to see if this helps
AIC decreases slightly only, I think it needs more fix
```{r}
xreg <- cbind(Temp=data$Temperature,DewPoint=data$`Dew Point`,SurfaceAlbedoLevel=data$SurfaceAlbedoLevel,WindDir =data$`Wind Direction`,WindSpeed=data$`Wind Speed`)

model1_predictor <- auto.arima(data$GHI,xreg = xreg)
model1_predictor
```

# model assesement
bad!!! because acf on residual does not dies down definitely
```{r}
checkresiduals(model1_predictor)
acf(residuals(model1_predictor))
pacf(residuals(model1_predictor))
```

# accuracy
#forecast looks better because temperature captures seasonality possibly
# ignore <0 prediction, it should be considred as 0.
```{r}
datahead <- data[1:8040,]
datatail <- data[8040:8760,]
xreg <- cbind(Temp=datahead$Temperature,DewPoint=datahead$`Dew Point`,SurfaceAlbedoLevel=datahead$SurfaceAlbedoLevel,WindDir =datahead$`Wind Direction`,WindSpeed=datahead$`Wind Speed`)

refit_predictor <- Arima(GHIhead, order=c(4,1,2),xreg = xreg)
checkresiduals(refit_predictor,lag.max=100)
arimapredictorfcast <- forecast(refit_predictor,
  xreg = cbind(Temp=datatail$Temperature,DewPoint=datatail$`Dew Point`,SurfaceAlbedoLevel=datatail$SurfaceAlbedoLevel,WindDir=datatail$`Wind Direction`,WindSpeed=datatail$`Wind Speed`
    ))
autoplot(GHIhead) +
  autolayer(arimapredictorfcast, series="forecast", PI=FALSE)

arimapredictorfcast %>% autoplot(ylab="Hourly GHI")
accuracy(arimapredictorfcast,GHItail)


```
terrible still, but normality assumption improved
stochastic trends, d=1


# I think I need seasonality, I don't know how https://hal.archives-ouvertes.fr/hal-02151290/document this only do it with arma(29,0)

#seasonally differencing
```{r}
# tsDiff24
tsDiff24 %>%ggtsdisplay(lag.max=100)
acf(tsDiff24,lag.max=100)
pacf(tsDiff24,lag.max=100)
```
```{r}
seasonaldiffARIMA2 <- Arima(tsGHI,order=c(1,1,3),seasonal=list(order=c(1,1,1),period=24))
seasonaldiffARIMA2
checkresiduals(seasonaldiffARIMA2,lag=80)
acf(seasonaldiffARIMA2$residuals,lag.max=80)
pacf(seasonaldiffARIMA2$residuals,lag.max=80)
```

```{r}
seasonaldiffARIMA1 <- Arima(tsGHI,order=c(4,1,3),seasonal=list(order=c(1,1,1),period=24))
seasonaldiffARIMA1
checkresiduals(seasonaldiffARIMA1,lag.max=100)
acf(seasonaldiffARIMA1$residuals,lag.max=80)
pacf(seasonaldiffARIMA1$residuals,lag.max=80)
```

```{r}
seasonaldiffARIMA2 <- Arima(tsGHI,order=c(5,1,3),seasonal=list(order=c(1,1,1),period=24))
seasonaldiffARIMA2
checkresiduals(seasonaldiffARIMA2,lag=80)
acf(seasonaldiffARIMA2$residuals,lag.max=80)
pacf(seasonaldiffARIMA2$residuals,lag.max=80)
```

```{r}
seasonaldiffARIMA2 <- Arima(tsGHI,order=c(7,1,3),seasonal=list(order=c(1,1,1),period=24))
seasonaldiffARIMA2
checkresiduals(seasonaldiffARIMA2,lag=80)
acf(seasonaldiffARIMA2$residuals,lag.max=80)
pacf(seasonaldiffARIMA2$residuals,lag.max=80)
```
#forecasting
```{r}
seasonalarimamodel_refit <- Arima(GHIhead,order=c(4,1,3),seasonal=list(order=c(1,1,1),period=24))
seasonalarimamodel_refit
checkresiduals(seasonalarimamodel_refit,lag.max=100)
```
```{r}
seasonalarimafcast <- forecast(seasonalarimamodel_refit,h=30*24)
autoplot(GHIhead) +
  autolayer(seasonalarimafcast, series="forecast", PI=FALSE)

seasonalarimafcast %>% autoplot()
accuracy(seasonalarimafcast,GHItail)
```

```{r}
dayahead_train <- window(mstsGHI,start=c(1,1),end=c(1,8736))
dayahead_test <- window(mstsGHI,start=c(1,8736),end=c(1,8760))
dayaheadGHI_refit <- Arima(dayahead_train,order=c(4,1,3),seasonal=list(order=c(1,1,1),period=24))
# mstsGHI_refit<-mstsGHI_refit["model"]$model
dayahead_fcast <-forecast(dayaheadGHI_refit,h=24)
autoplot(dayahead_train) +
  autolayer(dayahead_fcast, series="forecast", PI=FALSE)
dayahead_fcast %>% autoplot()+theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#+scale_x_date(limits =  as.Date(c("01-01", "12-31"),"%m-%d"))#,minor_breaks = as.Date(c("01-01", "04-01",\"08-01", "12-01"),"%m-%d"))

accuracy(dayahead_fcast,dayahead_test)

dayahead <- data.frame(hour=seq(0,23,1),ghi=dayahead_fcast$mean,ghi_actual = tsGHI[8737:8760])

ggplot(dayahead,aes(x=hour,y=ghi,color="predicted"))+geom_point()+geom_point(aes(x=hour,y=ghi_actual,color="expected"))+geom_smooth()+labs(title="One-day-ahead forecast for Dec.31st",y="Hourly GHI")
```



time series decomposition


```{r}
mstsGHI<-msts(data$GHI,seasonal.periods = c(24,8760/12))#first is the daily seasonality, then the monthly: period1 =24 hours, period2 = 24*30 =8760/12
mstlGHI <- mstl(mstsGHI)
mstlGHI %>% autoplot()
mstsGHI_train <- window(mstsGHI,start=c(1,1),end=c(1,8040))
mstlGHI_train <- mstl(mstsGHI_train)
mstlGHI_train %>% autoplot()
mstsGHI_test <- window(mstsGHI,start=c(1,8040),end=c(1,8760))

adjusted<- mstlGHI_train%>%seasadj()
adjustedDiff1 <- diff(adjusted)
acf(adjustedDiff1,lag.max=80)
pacf(adjustedDiff1,lag.max=80)

```
```{r}
adjusted_model <- auto.arima(adjusted)
checkresiduals(adjustedModel,lag.max=150)
```

```{r}
adjustedModel <- Arima(adjusted,order=c(2,1,2))
adjustedModel

```

```{r}
seasonalSTL0<-stlm(mstsGHI_train,modelfunction=Arima,order=c(5,1,2))#,seasonal=list(order=c(2,0,0),period=24))
seasonalSTLmodel0<-seasonalSTL0$model
seasonalSTLmodel0
checkresiduals(seasonalSTLmodel0,lag.max=150)
acf(residuals(seasonalSTLmodel0),lag.max=100)
AIC(seasonalSTL0$model)
ms_fcast <-forecast(seasonalSTL0,h=720)
ms_fcast%>%autoplot()
accuracy(ms_fcast,mstsGHI_test)
```



```{r}
monthahead <- data.frame(hour=seq(1,720,1),ghi=ms_fcast$mean,ghi_actual = mstsGHI[8041:8760])

ggplot(monthahead,aes(x=hour,y=ghi,color="predicted"))+geom_point()+geom_point(aes(x=hour,y=ghi_actual,color="expected"))+geom_smooth()+labs(title="One-day-ahead forecast for Dec.31st",y="Hourly GHI")
```
#accuracy

```{r}
dayahead_train <- window(mstsGHI,start=c(1,1),end=c(1,8592))
dayahead_test <- window(mstsGHI,start=c(1,8592),end=c(1,8760))
dayaheadGHI_refit <- stlm(dayahead_train,modelfunction=Arima,order=c(5,1,2))
# mstsGHI_refit<-mstsGHI_refit["model"]$model
dayahead_fcast <-forecast(dayaheadGHI_refit,h=24*7)
autoplot(dayahead_train) +
  autolayer(dayahead_fcast, series="forecast", PI=FALSE)
dayahead_fcast %>% autoplot()+theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
#+scale_x_date(limits =  as.Date(c("01-01", "12-31"),"%m-%d"))#,minor_breaks = as.Date(c("01-01", "04-01",\"08-01", "12-01"),"%m-%d"))

accuracy(dayahead_fcast,dayahead_test)

dayahead <- data.frame(hour=seq(1,24*7,1),ghi=dayahead_fcast$mean,ghi_actual = tsGHI[8593:8760])
dayahead$ghi[dayahead$ghi<0] <- 0
ggplot(dayahead,aes(x=hour,y=ghi,color="predicted"))+geom_line(lwd=1.5)+geom_line(aes(x=hour,y=ghi_actual,color="actual"),lwd=1)+labs(title="One-week-ahead forecast for Dec.24th-31st",y="Hourly GHI")#+geom_smooth()
```

```{r}
seasonalmodel2<-stlm(mstsGHI_train,modelfunction=Arima,order=c(4,1,2))
modelOnAdjusted2<-seasonalmodel2["model"]$model
modelOnAdjusted2
checkresiduals(seasonalmodel2)
acf(residuals(seasonalmodel2),lag.max=80)+theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())
pacf(residuals(seasonalmodel2),lag.max=80)+theme(axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())

forecast(seasonalmodel2,h=720)%>%autoplot()
```